---
title: "week3_VibigaVinotharajah"
output: html_document
date: '2022-05-05'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
```

```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
```


```{r}
caro_df <- read.csv("caro60.csv")
class(caro_df)
```

Step a) Specify the temporal window v (n+/-3) 6 positions:
1. pos[n-3] to pos[n]
2. pos[n-2] to pos[n]
3. pos[n-1] to pos[n]
4. pos[n] to pos[n+1]
5. pos[n] to pos[n+2]
6. pos[n] to pos[n+3]

Step b):
```{r}
caro <- caro_df %>% 
  mutate(
    nMinus3 = sqrt((lag(E,3)-E)^2+(lag(N,3)-N)^2),
    nMinus2 = sqrt((lag(E,2)-E)^2+(lag(N,2)-N)^2),
    nMinus1 = sqrt((lag(E,1)-E)^2+(lag(N,1)-N)^2),
    nPlus1  = sqrt((E-lead(E,1))^2+(N-lead(N,1))^2),
    nPlus2  = sqrt((E-lead(E,2))^2+(N-lead(N,2))^2),
    nPlus3  = sqrt((E-lead(E,3))^2+(N-lead(N,2))^2)
  )
```

```{r}
caro_df <- caro %>%
  rowwise() %>%
  mutate(
    stepMean = mean(c(nMinus3,nMinus2,nMinus1,nPlus1,nPlus2,nPlus3))
  ) %>%
  ungroup() 

```

Task 2:
```{r}
summary(caro_df$stepMean)
```

```{r}
p1 <- ggplot(caro_df, aes(x=stepMean))+
  geom_histogram(color="darkblue", fill="lightblue", bins=50)+
  ggtitle("Histogram of the Step Length from Caro ")+
  labs(x="Step Length in meters", y= "Count")

p1
```

```{r}
p2 <- ggplot(caro_df, aes(x= stepMean, y=TierName))+
  geom_boxplot()+
  labs(y="", x="Step Length in meters", title= "Boxplot of the Step Length from Caro")

p2

mean(caro_df$stepMean, na.rm=T)
```


```{r}
caro_df <- caro_df %>% 
  ungroup() %>%
  mutate(static = stepMean < mean(stepMean, na.rm = TRUE))


caro_filter <- caro_df %>%
  filter(!static)

caro_filter%>%
  ggplot(aes(E, N))  +
  geom_path() +
  geom_point() +
  coord_fixed() +
  theme(legend.position = "bottom")
```

Task 3:
```{r}
ggplot(caro_df, aes(x=E, y=N, colour= static ))+
  geom_path()+
  geom_point()+
  coord_equal()+
  ggtitle("Visualization of segmented trajectories")
```

Task 4:
```{r}
rle_id <- function(vec){
  x <- rle(vec)$lengths
  as.factor(rep(seq_along(x), times=x))
  }
```

```{r}
caro60 <- caro_df %>%
  mutate(segment_id = rle_id(static))
```

```{r}
#caro60_1 <- as_tibble(caro60)
#caro60 %>% filter(segment_id > 5)
caro_df <-caro60 %>% 
  slice(62:200) %>% 
    glimpse()
```


```{r}
ggplot(caro60, aes(x=E, y=N, colour= segment_id ))+
  geom_path()+
  geom_point()+
  coord_equal()+
  labs(title="Moving segments coloured by segment ID",
subtitle= "All segments (uncleaned)")
```

```{r}
ggplot(caro_df, aes(x=E, y=N, colour= segment_id ))+
  geom_path()+
  geom_point()+
  coord_equal()+
  labs("Moving segments coloured by segment ID",
subtitle="Long segments (removed segments <5)")
```


```{r}
ggplot(caro60, aes(x=E, y=N, colour= segment_id  ))+
  geom_path()+
  geom_point()+
  coord_equal()+
  ggtitle("Moving segments coloured by segment ID\n Long segment (removed segments <5")
```



Task 5:
```{r}
ped <- read.csv("pedestrian.csv")
class(ped)
```



```{r}
ped1 <- group_by(ped, TrajID)
ped1 <- mutate(ped,timelag = as.numeric(difftime(lead(DatetimeUTC),DatetimeUTC, units = "secs")))
```

```{r}
ped1 <- mutate(ped1,steplength = (sqrt((E-lead(E))^2 + (N-lead(N))^2)))
```

```{r}
ped1 <- mutate(ped1, speed = ((steplength/timelag)))
```


```{r}
summary(ped1)
length(ped1)
```

```{r}
ggplot(ped, aes(x=E, y=N, colour= factor(TrajID))) +
  geom_point() +
  facet_wrap("TrajID") +
  ggtitle("Visual comparison of the 6 trajectories\nEach subplot highlights a trajectory")+
  theme(legend.position="none")

```

```{r}
ped %>%
  group_by(TrajID) %>%
  summarise(sample_size = n())
```


In trajectory 5, there is an outlier. Trajectory 4 has a different shape and trajectory 3 pedestrian moved reverse.The point density between the sic trajectories is different, which shows that there are different speeds between the trajectories.

Task 6: 
```{r}
library(SimilarityMeasures)
```

```{r}
ped <- select(ped, -DatetimeUTC)

traj1 <- ped %>%
  filter(TrajID ==1) %>%
  as.matrix()
traj2 <- ped %>%
  filter(TrajID ==2) %>%
  as.matrix()
traj3 <- ped %>%
  filter(TrajID ==3) %>%
  as.matrix()
traj4 <- ped %>%
  filter(TrajID ==4) %>%
  as.matrix()
traj5 <- ped %>%
  filter(TrajID ==5) %>%
  as.matrix()
traj6 <- ped %>%
  filter(TrajID ==6) %>%
  as.matrix()
```

```{r}
DTW_1 = c(
         DTW(traj1, traj2, 4),
         DTW(traj1, traj3, 4),
         DTW(traj1, traj4, 4),
         DTW(traj1, traj5, 4),
         DTW(traj1, traj6, 4))

EditDist_1 = c(EditDist(traj1, traj2, 2),
               EditDist(traj1, traj3, 2),
               EditDist(traj1, traj4, 2),
               EditDist(traj1, traj5, 2),
               EditDist(traj1, traj6, 2))

```

```{r}
Frechet_1 = c(Frechet(traj1, traj2),
              Frechet(traj1, traj3),
              Frechet(traj1, traj4),
              Frechet(traj1, traj5),
              Frechet(traj1, traj6))

LCSS_1 = c(LCSS(traj1, traj2, 2, 2),
           LCSS(traj1, traj3, 2, 2),
           LCSS(traj1, traj4, 2, 2),
           LCSS(traj1, traj5, 2, 2),
           LCSS(traj1, traj6, 2, 2))
```

```{r}

my_measures <- data.frame(
  traj = 2:6,
  DTW = DTW_1,
  EditDist = EditDist_1,
  Frechet = Frechet_1,
  LCSS= LCSS_1
)

my_measures_long <- my_measures %>%
  pivot_longer(-traj) 
#"lengthens" data, increasing the number of rows and decreasing the number of columns.

ggplot(my_measures_long, aes(traj, value, fill = as.factor(traj)))+
  geom_col() +
  facet_wrap(~name, scales = "free_y") +
  theme(legend.position = "none") +
  labs(title="Computed similarities using different measures\n between trajectory 1 to all other trajectories")

```





